# Setup
```{r}
library(vcfR)
library(ggVennDiagram)
```


# Participants
## Ven diagrams
```{r}
# The 45 family study
participants_45 <-  get_pheno_data_wide() %>% 
  pull(sourceId) %>% 
  unique() %>% 
  list() %>% set_names("45")

# clinical data
participants_clinical <- get_clinical_data() %>% 
  pull(sourceId) %>% 
  unique() %>% 
  list() %>% 
  set_names("clinical")

# Anette data
participants_anette <- read_csv("data-raw/final_pheno.txt") %>% 
  pull(overlap) %>% 
  list() %>% 
  set_names("anette")

# the RNA seq data
participants_rna_seq <- make_salmon_metadata() %>% 
  pull(sourceId) %>% 
  unique() %>% 
  list() %>% 
  set_names("rna_seq")

# result
participants_per_project_w_rna_seq <- c(participants_45,
                                        participants_rna_seq, 
                                        participants_clinical,
                                        participants_anette)
ggVennDiagram(participants_per_project_w_rna_seq)

participants_anette_rna_seq <- c(participants_anette, 
                                 participants_rna_seq,
                                 participants_clinical)

ggVennDiagram(participants_anette_rna_seq)
process_region_data(Venn(participants_per_project_w_rna_seq)) %>% filter(name=="rna_seq..clinical") %>% pull(item) %>% set_names("Not in Phenomics but present in other data")
process_region_data(Venn(participants_per_project_w_rna_seq)) %>% filter(name=="rna_seq") %>% pull(item) %>% set_names("Not present in Phenomcis or other data")
```
## Lists
```{r}
setdiff(participants_rna_seq$rna_seq, intersect(participants_per_project$`45`, participants_rna_seq$rna_seq))
setdiff(participants_rna_seq$rna_seq, intersect(participants_per_project$`45`, participants_rna_seq$rna_seq))
```

## Conclusion: 4/5 of the missing samples in the Phenomics data are found in the clinical data. So they must exist somewhere.

# Duplicate rows? 
```{r}
pheno_data %>% distinct(sourceId, phenoVar_id, phenoVariable, result)
```

# VennDiagrams of shared genetic data
```{r}
vcf_files <- list.files(path = "data-raw/vcf_files", 
                        pattern = ".*dose.vcf.gz$", full.names = T)

vcfr_objects <- map(vcf_files, read.vcfR)
```

# Clinical data vs phenomics data
```{r}
pheno_data_wide <- get_pheno_data_wide()
clinical_data <- get_clinical_data()

df <- inner_join(pheno_data_wide, clinical_data, by=c("sourceId"="ID")) %>% 
  ggsankey::make_long(`Far (1) mor (2) syskon (3)barn (4) med T2D`, fam_var)

ggplot(df, aes(x = x, 
               next_x = next_x, 
               node = node, 
               next_node = next_node,
               fill = factor(node),
               label = node)) +
  ggsankey::geom_sankey() +
  ggsankey::geom_sankey_label() +
  ggsankey::theme_sankey(base_size = 16)
```

# Old
## VennDiagrams of shared phenotype variables
```{r}
phenotypes_per_project <- map(projects, ~pheno_data %>% 
                                filter(projectId==.x) %>% 
                                pull(phenoVariable) %>% 
                                unique())

ggvenn(phenotypes_per_project)
```

## Conclusion: All source IDs (and thereby metadata) are contained in project 45

