# Correlation between phenotypes
```{r}
clinical_data_subset <- get_clinical_data() %>% 
  select(HOMA_B_2.88, 
         BIGTT_AIR_t_0_60_120, 
         Insulinogenic_index, 
         HOMA_IR_156, 
         BIGG_SI_t_0_60_120, 
         Homa_is,  
         BMI,
         AGE,
         WH) %>% 
  drop_na()
```

## Matrix
```{r, fig.height=10, fig.width=10}
library(ggcorrplot)

cor_data <- clinical_data_subset %>% 
  cor()

p_mat <- clinical_data_subset %>% cor_pmat()


cor_data %>% ggcorrplot(lab = TRUE,insig = "blank", p.mat = p_mat)
```
## Plot
```{r}
get_clinical_data() %>% 
  ggplot(aes(x=WH, y=BIGG_SI_t_0_60_120)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~SEX)
```



# Oral Gluose Test
```{r}
plot_oral_glucose_data()
```

# Raw CPM Visualizations
```{r}
# muscle
pheno_data_w_RNA <- get_pheno_data_w_RNA("2022-09-19-15.52_analysis_muscle_data/")
# fat
# pheno_data_w_RNA <- get_pheno_data_w_RNA("2022-09-19-15.46_analysis_fat_data/")
```

## BMI and leptin
```{r}
pheno_data_w_RNA %>% 
  ggplot(aes(x=BMI, y=LEP)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~SEX)

pheno_data_w_RNA %>% 
  ggplot(aes(x=LEP, y=leptin)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~SEX)

pheno_data_w_RNA %>% 
  ggplot(aes(x=BMI, y=leptin)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~SEX)
```

## Gene CPMs and various IR/IS measures
### All type 2 diabetes associated genes versus IR/IS measures
```{r, message=FALSE}
# IR_IS_measures <- c("HOMA_IS_156","BIGG_SI_t_0_60_120","WBSI")
IR_IS_measures <- c("BIGG_SI_t_0_60_120")
genes <- get_interesting_genes()



# Pivot RNA data to long format according to various IR/IS measures
pheno_data_w_RNA_measure_long <- pheno_data_w_RNA %>%
  # mutate(HOMA_IS_156=1/HOMA_IR_156) %>%
  pivot_longer(cols=all_of(IR_IS_measures),
               names_to = "IS_name",
               values_to = "IS_value")

# Pivot again to even longer
pheno_data_w_RNA_measure_genes_long <- pheno_data_w_RNA_measure_long %>%
  pivot_longer(cols=any_of(genes),
               names_to="Gene_name",
               values_to="Gene_CPM")

# for (i in seq(1,length(unique(pheno_data_w_RNA_measure_genes_long$Gene_name)-10), 10)) {
# print(i)
my_plot <- pheno_data_w_RNA_measure_genes_long %>%
  ggplot(aes(x=Gene_CPM, y=IS_value)) +
  geom_point(size=0.1) +
  scale_y_continuous(expand = expansion(mult=c(0,0.5))) +
  geom_smooth() +
  ggforce::facet_wrap_paginate(facets = vars(Gene_name), nrow = 2, ncol = 2,
                               scales = "free")


for (i in 1:ggforce::n_pages(my_plot)) {
  print(i)
  page_plot <- my_plot + ggforce::facet_wrap_paginate(facets = vars(Gene_name), 
                                                      nrow = 2,
                                                      ncol = 2,
                                                      scales = "free",
                                                      page=i)
  
  plot(page_plot)
}
```

### Exercise related genes in healthy adults vs IS measure
```{r}
# genes <- get_long_term_exercise_induced_genes()
genes <- c("MYH1","SH3RF2", "MSTN","MYH2")
IR_IS_measures <- c("BIGG_SI_t_0_60_120")

pheno_data_w_RNA_measure_long <- pheno_data_w_RNA %>%
  pivot_longer(cols=all_of(IR_IS_measures),
               names_to = "IS_name",
               values_to = "IS_value")

# Pivot again to even longer
pheno_data_w_RNA_measure_genes_long <- pheno_data_w_RNA_measure_long %>%
  pivot_longer(cols=any_of(genes),
               names_to="Gene_name",
               values_to="Gene_CPM")

pheno_data_w_RNA_measure_genes_long %>%
  ggplot(aes(x=Gene_CPM, y=IS_value)) +
  geom_point(size=0.1) +
  scale_y_continuous(expand = expansion(mult=c(0,0.5))) +
  # geom_smooth() +
  # facet_grid(rows = vars(SEX),
  #           cols = vars(Gene_name),
  #            scales = "free")
  facet_wrap(facets=vars(Gene_name),
             scales="free", ncol=2)


```

```{r}
pheno_data_w_RNA_fat_long %>% 
  ggplot(aes(x=SLC2A4, y=IS_value)) +
  geom_point() +
  scale_y_continuous(expand = expansion(mult=c(0,0.5))) +
  facet_wrap(~IS_name, scales="free") +
  xlab("GLUT4 gene expression (CPM)") +
  ylab("Insulin sensitivity index value") +
  geom_smooth(method = "lm")
```

### Fitted vs raw counts
```{r}
fit_all <- readRDS("2022-09-19-15.46_analysis_fat_data/fit_all.RDS")
fit_all$coefficients %>% as_tibble(rownames = "Gene") %>% filter(Gene=="SLC2A4")

fitted <- fit_all$fitted.values %>% as_tibble(rownames = "Gene") %>% filter(Gene=="SLC2A4") %>% pivot_longer(cols=2:last_col())

raw <- fit_all$counts %>% as_tibble(rownames = "Gene") %>% filter(Gene=="SLC2A4") %>% pivot_longer(cols=2:last_col())

combined <- inner_join(raw, fitted, by="name")
combined %>% ggplot(aes(x=value.x, y=value.y)) +
  geom_point()
```

## PRKCA and HOMA_IR_156
```{r}
pheno_data_w_RNA_muscle %>% 
  ggplot(aes(x=HOMA_IR_156, y=PRKCA)) +
  geom_point() +
  geom_smooth(method = "lm")
```


# Relatedness matrix PCA
```{r, fig.width=8, fig.height=14}
plot_PCA()
```

# Correlation between waist-hip-ratio and BIGtt_S in WH-risk high group
## Setup
```{r}
combined_pheno_data <- get_combined_pheno_data()
combined_pheno_data_WH_risk_high <- combined_pheno_data %>% 
  filter(sourceId!="641-9") %>% 
  filter(WH_risk=="High")
```

```{r}
combined_pheno_data_WH_risk_high %>% distinct(sourceId, .keep_all = T) %>% 
  ggplot(aes(x=BIGG_SI_t_0_60_120, y=WH)) +
  geom_point() +
  facet_wrap(~SEX)

combined_pheno_data_WH_risk_high %>% distinct(sourceId, .keep_all = T) %>% 
  select(~SEX)
```

