# Models and running
```{r}
# Insulin sensitivity and secretion model
# interesting_formula <- "~WBSI + CIR + WBSI:CIR + SEX + AGE"
# interesting_coefficients <- c("WBSI","CIR","WBSI.CIR")

# Insulin sensitivity/secretion and WH model
interesting_formula <- "~WBSI + CIR + WBSI:CIR + WH + SEX + AGE"
interesting_coefficients <- c("WBSI","CIR","WBSI.CIR", "WH")

# Insulin sensitivity model only
# interesting_formula <- "~WBSI + SEX + AGE"
# interesting_coefficients <- c("WBSI")

results <- run_full_analysis(tissues = c("muscle","fat"),
                             formulas = interesting_formula)

# results <- readRDS("results_muscle_fat_DI_SEX_AGE_19_13_26.RDS")
# results <- readRDS("results_muscle_fat_WBSI_CIR_WBSI:CIR_SEX_AGE_19_15_09.RDS")
results <- readRDS("results_muscle_fat_WBSI_CIR_WBSI:CIR_WH_SEX_AGE_19_16_40.RDS")

```


# Comparison of genes in fat and muscle
## Based on median CPMS
There is some correlation between CPM in fat and muscle within the individuals, but it is not crazy much, i.e. ~0.5.
```{r}
muscle_long <- pivot_longer(data = results$muscle$DATA$samples,
                            cols = A1BG:last_col(),
                            names_to="Gene",
                            values_to="CPM_muscle") %>% 
  select(sourceId, Gene, CPM_muscle)

fat_long <- pivot_longer(data = results$fat$DATA$samples,
                         cols = A1BG:last_col(),
                         names_to="Gene",
                         values_to="CPM_fat") %>% 
  select(sourceId, Gene, CPM_fat)


plotting_data <- full_join(muscle_long, fat_long) %>% 
  group_by(Gene) %>% 
  summarise(med_CPM_muscle=median(CPM_muscle), med_CPM_fat=median(CPM_fat))

corr_val <- cor(plotting_data$med_CPM_muscle, plotting_data$med_CPM_fat) %>% 
  signif(2)

plotting_data %>% 
  ggplot(aes(x=med_CPM_muscle, y=med_CPM_fat)) +
  geom_point(alpha=0.5, size=0.1) +
  xlim(c(0,500)) +
  ylim(c(0,500)) +
  geom_smooth(method = "lm") +
  ggtitle(str_glue("Correlation bewteen median fat and muscle CPMs: {corr_val}"))



```
## Based on log-fold change from interesting coefficientresistance
```{r, fig.width=15, fig.height=15}
for (coefficient in interesting_coefficients) {
  
  plotting_data_log_fold_change <- full_join(results$muscle[[interesting_formula]]$DGE[[coefficient]],
                                             results$fat[[interesting_formula]]$DGE[[coefficient]],
                                             by="Name",
                                             suffix = c("_muscle","_fat"))
  
  plot <- plotting_data_log_fold_change %>% 
    filter(FDR_fat<0.05 & FDR_muscle<0.05) %>% 
    ggplot(aes(x=logFC_fat,
               y=logFC_muscle,
               label=Name,
               size=-log10(FDR_fat),
               color=-log10(FDR_muscle))) +
    geom_point(alpha=0.5) +
    geom_abline(slope = 1) +
    ggtitle(coefficient)
  
  plotly_plot <- plotly::ggplotly(plot)
  
  print(plotly_plot)
}
```

# Euler plots
## DE 
### Bewteen tissues
Many more differentially expressed genes in fat compared to muscle. ~50 % of differentially expressed genes in muscle are also differentially expressed in fat.
```{r}
for (coefficient in interesting_coefficients) {
  
  DE_input_list <- list("muscle"=results$muscle[[interesting_formula]]$DGE[[coefficient]] %>% get_sig_entities_from_df(type = "DE"),
                        "fat"=results$fat[[interesting_formula]]$DGE[[coefficient]] %>% get_sig_entities_from_df(type = "DE"))
  
  DE_fit <- euler(DE_input_list)
  my_plot <- plot(DE_fit, 
                  quantities = TRUE,
                  labels = list(font = 4),
                  main=str_glue("{coefficient}_DE"))
  
  print(my_plot)
}
```

## Between coefifcients
```{r}
for (tissue in tissues) {
  DE_input_list <- map(.x = interesting_coefficients %>% set_names(), 
                       .f = function(coefficient) {
                         sig_entities <- results[[tissue]][[interesting_formula]]$DGE[[coefficient]] %>% get_sig_entities_from_df(type = "DE", genes_type = "NCBI")
                       }
  )
  
  DE_fit <- euler(DE_input_list)
  
  my_plot <- plot(DE_fit, 
                  quantities = TRUE,
                  labels = list(font = 4),
                  main=str_glue("{tissue}_DE"))
  
  print(my_plot)
}
```

## DE with gene categories
Half of the exercise genes are differentially regulated in fat, but only one in muscle.
```{r}
exercise_induced_genes_list <- list("exercise_genes"=get_long_term_exercise_induced_genes())
insulin_sensitivty_genes <- get_interesting_genes_df() %>% 
  split(f = as.factor(.$GROUP)) %>% 
  map(~pull(., NAME))

DE_fit_w_other_genes <- euler(combinations = c(exercise_induced_genes_list, DE_input_list))#,
# control=list("extraopt_control"=list("max.call"=1e3)))
plot(DE_fit_w_other_genes, 
     quantities = TRUE,
     labels = list(font = 4),
     main="DGE")
```

## GO
### Between tissues
We have most GO-terms in fat compared to muscle. More than 50 % of GO-terms in muscle are also differentially enriched in fat.
```{r}
for (coefficient in interesting_coefficients) {
  
  GO_input_list <- list("muscle"=results$muscle[[interesting_formula]]$GO[[coefficient]] %>% get_sig_entities_from_df(type = "GO"),
                        "fat"=results$fat[[interesting_formula]]$GO[[coefficient]] %>% get_sig_entities_from_df(type = "GO"))
  
  GO_fit <- euler(GO_input_list)
  my_plot <- plot(GO_fit, 
                  quantities = TRUE,
                  labels = list(font = 4),
                  main=str_glue("{coefficient}_GO"))
  
  print(my_plot)
}
```
### Between coefficients
```{r}
for (tissue in tissues) {
  GO_input_list <- map(.x = interesting_coefficients %>% set_names(), 
                       .f = function(coefficient) {
                         sig_entities <- results[[tissue]][[interesting_formula]]$GO[[coefficient]] %>% get_sig_entities_from_df(type = "GO")
                       }
  )
  
  GO_fit <- euler(GO_input_list)
  
  my_plot <- plot(GO_fit, 
                  quantities = TRUE,
                  labels = list(font = 4),
                  main=str_glue("{tissue}_GO"))
  
  print(my_plot)
}
```


# Comparing the glucose tolerance groups based on clinical data
## All
The glucose tolerance class 3 are more obese, and show many signs of diabetes.
```{r, fig.height = 10, fig.width=7, warning=FALSE, message=FALSE}
clinical_data<- get_clinical_data()

measures <- list("insulin_secretion_measures" = c("CIR", "HOMA_B_2.88","Insulinogenic_index","BIGTT_AIR_t_0_60_120"),
                 "insulin_sensitivty_measures" = c("WBSI","HOMA_IR_156","BIGG_SI_t_0_60_120"),
                 "disposition_measures" =c("HBA", "DI"),
                 "anthropologic_measures"=c("WAIST","WH", "WEIGHT","BMI", "HIP", "HEIGHT", "AGE"))


imap(measures, 
     ~{plot_data <- clinical_data %>% 
       pivot_longer(cols = all_of(.x))
     
     box_plot <- plot_data %>% 
       ggplot(aes(color=Glu_tol_3_class, x=Glu_tol_3_class, y=value, label=sourceId)) +
       geom_point(size=0.1) +
       geom_boxplot() +
       geom_smooth(aes(group=1)) +
       facet_wrap(facets = vars(name), 
                  scales = "free") +
       ggtitle(.y) 
     
     corr_plot <- clinical_data %>% 
       select(all_of(.x)) %>% 
       drop_na() %>% 
       cor() %>%
       ggcorrplot::ggcorrplot(lab = TRUE)
     
     cowplot::plot_grid(box_plot, 
                        corr_plot,
                        ncol=1)
     }
)

non_anthro_measures <- measures[c("insulin_secretion_measures",
                                  "insulin_sensitivty_measures",
                                  "disposition_measures")] %>% unlist() %>% unname()
clinical_data %>% 
  select(all_of(non_anthro_measures)) %>% 
  drop_na() %>% 
  cor() %>%
  ggcorrplot::ggcorrplot(lab = TRUE)

```
# Dot plots showing outliers
```{r, fig.width=10, fig.height=15}

clinical_data<- get_clinical_data()

plot_BMI <- clinical_data %>% 
  pivot_longer(cols = all_of(non_anthro_measures) & !(WH | BMI)) %>% 
  ggplot(aes(color=Glu_tol_3_class, x=BMI, y=value, label=sourceId)) +
  geom_point() +
  geom_smooth(aes(group=1), method="lm") +
  facet_wrap(facets = vars(name), 
             scales = "free", 
             ncol= 1) +
  theme(legend.position = "none")

# plotly::ggplotly(plot_BMI)

plot_WH_sex_specific <- clinical_data %>% 
  pivot_longer(cols = all_of(non_anthro_measures) & !(WH | BMI)) %>% 
  ggplot(aes(color=Glu_tol_3_class, x=WH, y=value, label=sourceId)) +
  geom_point() +
  geom_smooth(aes(group=1), method="lm") +
  facet_grid(rows = vars(name), 
             cols = vars(SEX),
             scales="free") +
  theme(legend.position = "none")
# plotly::ggplotly(plot_WH)


plot_WH_non_sex_specific <- clinical_data %>% 
  pivot_longer(cols = all_of(non_anthro_measures) & !(WH | BMI)) %>% 
  ggplot(aes(color=Glu_tol_3_class, x=WH, y=value, label=sourceId)) +
  geom_point() +
  geom_smooth(aes(group=1), method="lm") +
  facet_wrap(facets = vars(name),
             scales = "free",
             ncol = 1) +
  theme(legend.position = "none")
# plotly::ggplotly(plot_WH)


# comparison of BMI and WH
cowplot::plot_grid(plot_BMI, plot_WH_non_sex_specific, ncol = 2)

# Comparison of WH sex and non-sex specific
cowplot::plot_grid(plot_WH_sex_specific, plot_WH_non_sex_specific, ncol = 2, rel_widths = c(1.7,1))



```


# Tables
## DE
```{r}
for (tissue in tissues) {
  for (coefficient in interesting_coefficients) {
    my_table <- results[[tissue]][[interesting_formula]]$DGE[[coefficient]] %>% 
      filter(FDR<0.5) %>% 
      get_DE_datatable_updated(caption=str_glue(str_glue("{tissue}_{coefficient}")))
    
    print(my_table)
  }
}
```
## GO
```{r, fig.width=10, fig.height=20}
for (tissue in tissues) {
  for (coefficient in interesting_coefficients) {
    my_table <- results[[tissue]][[interesting_formula]]$GO[[coefficient]] %>% 
      filter(FDR<0.5) %>% 
      get_GO_datatable_updated(caption=str_glue(str_glue("{tissue}_{coefficient}")))
    
    print(my_table)
  }
}
```
# Volcano plots
## DE
```{r}
for (tissue in tissues) {
  for (coefficient in interesting_coefficients) {
    my_plot <- results[[tissue]][[interesting_formula]]$DGE[[coefficient]] %>% 
      ggplot_volcano_updated(only_sig = T, 
                             highlight_genes = c("MYH1", "MYH4"), 
                             title = str_glue("{tissue}_{coefficient}"))
    print(my_plot)
  }
}
```


